---
title: "Animating best place to stay in Edinburgh using the Yelp dataset"
author: "kimnewzealand"
date: "26 April 2018"
output: html_document
---

In this project, I would like to identify the best place to stay in Edinburgh given a preference for Japanese restaurants. 

I would like to visualise how much the number of people visiting restaurants changes by the location of the restaurant or the day of the week in Edinburgh.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
# Getting data
library(httr)
library(jsonlite)
# Manipulating strings
library(stringr)
# Manipulating date and time
library(lubridate)
# Mapping packages
library(maps)
library(ggthemes)
library(ggmap)
# devtools::install_github("dgrtwo/gganimate")
# devtools::install_github("yihui/animation")
library(gganimate)
library(ggalt)
```

## Load Data

The [Yelp datasets](https://www.kaggle.com/yelp-dataset/yelp-dataset) have been cleaned into csv formats from JSON and so they are easy to load with the read_csv function from the [readr](https://cran.r-project.org/web/packages/readr/index.html) R package.

These files have been downloaded into a local directory first to agree to the terms of use.

```{r load data}
# Read in the check in data
yelpcheck <- read_csv("yelp_checkin.csv")
# Read in the business data
yelpbus <- read_csv("yelp_business.csv")
```

The two data files loaded are:

**yelp_business.csv** includes locations and details of the  businesses by business_id. The location details are longitude and latitude which can be plotted with ggplot.  
**yelp_checkin.csv** includes number of check ins by day and time by business_id. This business_id can used as an id for joining the two datasets.

```{r clean data}
# Convert the hour column nane to whathour and character
names(yelpcheck)[3] <- "whathour"
yelpcheck$whathour <- as.character(yelpcheck$whathour)

# https://stat.ethz.ch/R-manual/R-devel/library/base/html/strptime.html

yelpcheck$weekday <- recode(yelpcheck$weekday, 
       "Sun"="Sunday",
       "Mon"="Monday",
       "Tue"="Tuesday",
       "Wed"="Wednesday",
       "Thu"="Thursday",
       "Fri"="Friday",
       "Sat"="Saturday")
```

## Filter data

Since we are only analysing the city Edinburgh, we will filter the dataset by the city.

```{r Filter by city}
# Filter by city Edinburgh
acity <- "Edinburgh"
yelpcity <- yelpbus %>% 
      filter(city==acity) 
# Check for outliers
range(yelpcity$longitude)
range(yelpcity$latitude)
plot(yelpcity$longitude,yelpcity$latitude)
# Remove the outliers
yelpcity <- yelpcity %>%
      filter(longitude > -5) %>%
      filter(latitude < 60)
```

## Aggregate datasets

Next we will aggregate the check-in data at a weekday level to get an initial visualisation of the checkins by weekday. Later we could change the aggregation to view the data on an hourly basis.

```{r aggregate}
# Calculate the sum of the checkins grouped by the business and weekday
yelpchecksum <- yelpcheck %>% 
      group_by(business_id,weekday) %>% 
      summarize_at( .vars = colnames(.)[4] , sum)
# Merge the city data and the sum of checkins
yelpcity <- merge(yelpcity,yelpchecksum, by ="business_id")
# Filter by a category, where each category is a list of category types separated by ;
acategory <- "Japanese"
yelpcitycat <- yelpcity %>% 
            filter(str_detect(categories, acategory))
yelpcitycat$weekday <- factor(yelpcitycat$weekday, levels=c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))
```

## Visualisations

First let's get the base map of the city Edinburgh and plot a ggplot object using facet wrap by the weekday. This will give us an indication of whether we have the right map size given our filtered data by restaurant type.

```{r staticmaps}
# Get the Edinburgh base map
edinmap <- get_map(location = c(lon = mean(yelpcitycat$longitude), lat = mean(yelpcitycat$latitude)), maptype = "roadmap", zoom = 14)
# Map using ggmap function and the theme void
ggmap(edinmap) + 
   yelpcitycat %>% 
      geom_point(mapping = aes(x = longitude, y = latitude, size = checkins, color = stars), alpha = 0.5) +
      scale_color_gradientn(colours = rainbow(5)) +
      theme_void() + 
      ggtitle("Edinburgh") +
      facet_grid(~weekday)
```
  
There appear to be a couple blue of 4 star restaurants and two purple 5 star restaurants.

Next create the [gganimate](https://www.rdocumentation.org/packages/gganimate/versions/0.1.0.9000/topics/gganimate) R package to animate checkins and ratings on a map. with the Old Town circled.

```{r gganimation, interval = 2}
oldtown <- yelpcity %>% filter(neighborhood=="Old Town")
# Create the ggplot object g 
g <- ggmap(edinmap) +  
      theme_void() + 
      ggtitle("Japanese Restaurants in Edinburgh") + 
      geom_encircle(aes(longitude, y = latitude),
                                     data = oldtown, size = 2, color = "orange")
# Use suppressWarnings to ignore warning messages which halt the plotting
suppressWarnings(g <-  g +
                       geom_point(data=yelpcitycat,mapping = aes(x = longitude, y = latitude, size = as.factor(stars), frame = weekday, color = checkins))) 
# gganimate the g object
gganimate(g)
gganimate(g, "edinb.gif")
```


In order to get the gganimate plot to work I had to install the github versions of the two packages gganimate and animation and use the suppressWarnings function as the warnings appeared to be halting the animation.

  
devtools::install_github("dgrtwo/gganimate")  
devtools::install_github("yihui/animation")




